Rotation and Lithium Confirmation of a 500 Parsec Halo for the Open Cluster NGC\,2516\footnote{This is paper 3 of the Cluster Difference Imaging Photometric Survey.}

### 2021-04-23 cluster_paper/ms.tex
update ms

Figure~\ref{fig:hr} shows the HR diagram in the space of absolute Gaia
$G$ magnitude as a function of observed $\bpmrp$ color.  Spectral
types are interpolated from the \citet{pecaut_mamajek_2013} table,
after accounting for reddening as described
below.\footnote{\url{http://www.pas.rochester.edu/~emamajek/EEM_dwarf_UBVIJHK_colors_Teff.txt},
version 2021.03.02.} Stars in the core appear consistent with having a
fixed age and metallicity, and varying mass.  The halo stars show a
similar sequence, though with greater scatter.  One possible
explanation for the additional scatter is that the halo is more
contaminated by field stars.  For instance, $\approx5$ red giants in
the halo must be field interlopers, because their isochronally implied
ages would be inconsistent with that of the cluster.

@@

\input{maintable_header.tex}
\begin{figure*}[t]
	\begin{center}
		\leavevmode
		\subfloat{
			\includegraphics[width=0.7\textwidth]{f3a.pdf}
		}

    \vspace{-0.5cm}
		\subfloat{
			\includegraphics[width=0.7\textwidth]{f3b.pdf}
		}
	\end{center}
	\vspace{-0.7cm}
  \caption{ {\bf TESS rotations periods and dereddened Gaia colors
  across the core and halo of \cn.} Sets $\mathcal{A}$ ({\it top})
  and $\mathcal{B}$ ({\it bottom}) undergo successive stages of
  automated cleaning (see Section~\ref{subsubsec:cluster}).  The Pleiades
  \citep[125\,Myr;][]{rebull_rotation_2016a} and Praesepe
  \citep[650\,Myr;][]{douglas_poking_2017} are shown for reference.
  Stars in the core and the halo of NGC\,2516 overlap with the
  Pleiades at $\bpmrpo$$<$1.2.  From $1.2<\bpmrpo<1.7$, some NGC\,2516
  members have longer rotation periods than Pleiades members. Rotation
  periods for comparison field stars are shown in
  Appendix~\ref{app:compstar}.
  \label{fig:rot}
	}
\end{figure*}
\begin{figure*}[t]
	\begin{center}
		\leavevmode
		\subfloat{
			\includegraphics[width=0.7\textwidth]{f3c.pdf}
		}

    \vspace{-0.5cm}
		\subfloat{
			\includegraphics[width=0.7\textwidth]{f3d.pdf}
		}
	\end{center}
	\vspace{-0.7cm}
  \caption{ {\bf Binarity indicators for the rotators in
  Figure~\ref{fig:rot}.} Sets $\mathcal{A}$ ({\it top}) and
  $\mathcal{B}$ ({\it bottom}) undergo the same successive stages of
  cleaning.  Astrometric binaries ($\mathrm{RUWE}>1.2$; 11\% of stars)
  appear in red; photometric binaries ($>$0.3\,mag from an empirical
  isochrone; 20\% of stars) are overplotted in orange.  
  \label{fig:binarity}
	}
\end{figure*}

We began our rotation analysis by considering all \nkinematic\
candidate members of \cn.  For each source, we retrieved all available
CDIPS light curves, on a per-sector basis.  This yielded \ncdips\
stars with at least one sector from a CDIPS light curve.  Each of
these stars is brighter than the CDIPS magnitude limit of
$Rp=16$\,mag.  \nkinematicrpltsixteen\ of the stars from the \cn\
source list had $Rp<16$.  The difference (65 stars) is due to 35 stars
from \citet{meingast_2021} which were not available at the time of the
CDIPS reductions, and 30 stars falling on the TESS chip gaps.  At the
distance of \cn, the $Rp=16$\,mag cutoff imposed during the CDIPS
processing corresponds roughly to $\bpmrpo$ of 2.2, or a spectral type
of $\approx$M2V.

We removed systematic trends shared between all light curves on each
CCD in each sector individually, and stitched together the resulting
light curves before searching for rotation-induced periodicity.
Details regarding our detrending approach are presented in
Appendix~\ref{app:detrending}.  After detrending, we proceeded with a
few cleaning steps: we masked 0.7 days at the beginning and end of
each spacecraft orbit, and ran a sliding standard-deviation rejection
window over the light curve, which removed any outlying points within
$\pm3\times$MAD of the median in each window.

We then measured the rotation period from the resulting light curve
using the periodogram implementations in \texttt{astrobase}, namely
the traditional Lomb-Scargle, and the Stellingwerf phase-dispersion
minimization algorithm
\citep{lomb_1976,stellingwerf_period_1978,scargle_studies_1982,stellingwerf_period_2011,bhatti_astrobase_2018}.
We used the CDIPS aperture radius that, based on theoretical
expectations, was expected to give the optimal balance between light
from the target star and background-light \citep{Sullivan_2015}.  This
typically resulted in a circular aperture radius of either 1 or 1.5 pixels.
We recorded the top five periodogram peaks from each method, and their
corresponding powers.  Finally, as a check on crowding, we recorded
the number of stars within the aperture with greater
brightness than the target star, and with brightness within 1.25 and
2.5 TESS magnitudes of the target star.

@@

\begin{figure*}[t]
	\begin{center}
		\leavevmode
			\includegraphics[width=0.9\textwidth]{f5a.pdf}
	\end{center}
	\vspace{-0.7cm}
  \caption{ {\bf Lithium in the core and halo of NGC\,2516.}
  Equivalent width of the 6708\,\AA\ doublet is shown versus
  dereddened color for all kinematic \cn\ members with Gaia-ESO
  ($N=452$) or GALAH ($N=107$) spectra available.  The GALAH spectra
  comprise slightly over half of the halo stars, due to the
  non-targeted selection function of that survey.  Field stars (gray
  points) are stars with detected
  transiting planets from Kepler 
  \citep{berger_identifying_2018}.  Stars in the Pleiades (125\,Myr)
  with Li detections are from \citet{bouvier_pleiades_lirot_2018};
  Praesepe detections (650\,Myr) are shown in violet
  \citep{soderblom_praesepe_1993}.  Points with ${\rm
  EW}\approx0\,$m\AA\ are non-detections.
  The Li measurements for both core and halo stars in \cn\ are
  consistent with a near-Pleiades age.
  {\bf LGB TODO: the increasing EWs for the Gaia-ESO M dwarfs must be erroneous. What's happening there?}
  \label{fig:lithiumcorehalo}
  }
\end{figure*}
\begin{figure*}[t]
	\begin{center}
		\leavevmode
		\includegraphics[width=0.9\textwidth]{f5b.pdf}
	\end{center}
	\vspace{-0.7cm}
  \caption{ {\bf Lithium and rotation in NGC\,2516.} Points show all
  kinematic \cn\ members with TESS rotation periods in the Gaia-ESO or
  GALAH surveys, and are colored by equivalent width.  Points with
  reported equivalent widths below 20\,m\AA\ (including
  non-detections) are shown with gray crosses.  The lithium-rotation
  correlation for the K-dwarfs has been observed in other young
  clusters, and is discussed in Section~\ref{subsec:lithium}.
	\label{fig:lithiumrot}
	}
\end{figure*}

@@

Gaia-ESO selected candidate \cn\ members to observe with the GIRAFFE
and UVES spectrographs based on previously reported literature members
and publicly available photometry.  Since the existence of the \cn\
halo was not known at the time of target selection, very few halo
stars are in the sample.  For \cn, \citet{randich_gaiaeso_2018}
reported stellar parameters (including lithium equivalent
widths and metallicities) for 796 stars that they considered possible
\cn\ members.  Cross-matching against our kinematic list of
\nkinematic\ candidate members by position and imposing a 0$\farcs$5
maximum separation limit yields 492 kinematic members for which a
Gaia-ESO spectrum was acquired and stellar parameters were reported.
15 of these (all with $\bpmrpo > 2.0$) are spurious matches based on
the Gaia color and GES effective temperature, and we remove them.
This yields 477 stars, of which 436 are in the core, and 41 are in the
halo.  The lop-sided ratio is due to the Gaia-ESO selection function.
We verified by directly querying the ESO archive that
\citet{randich_gaiaeso_2018} indeed included $>99\%$ of the available
Gaia-ESO spectra as part of their \cn\ sample, and so we opted to use
their results rather than repeating their analysis.

The GALAH DR3 target selection is discussed in depth by
\citet{buder_galah_2020}.  The relevant aspects for our analysis are
that the ``main'' survey targeted $12<V<14$ stars at
$\delta<+10^\circ$, provided the stars were at least ten degrees from
the galactic plane.  The survey is spatially
inhomogeneous.\footnote{See the footprint at
\url{https://www.galah-survey.org/news/announcing-galah-dr3}} Special
targeting for stars in the TESS southern continuous viewing zone, and
for known open cluster members was also performed.  We identified the
candidate \cn\ members for which spectra had been obtained by
searching the \texttt{GALAH\_DR3\_main\_allstar\_v1} catalog, after
excluding stars with the stellar parameter bit flags 1, 2, 3.  This
excludes spectra with unreliable broadening, low S/N, and unreliable
wavelength solutions (see Table~6 of \citealt{buder_galah_2020}).
Since our main focus is measuring equivalent widths for the
Li\,6708\,\AA\ doublet, these were the most relevant bit flags.  Of
our \nkinematic\ candidate \cn\ members, 107 had spectra in GALAH DR3.
51 were in the ``core''; 56 were in the ``halo''.  78 had ``finite''
lithium, i.e., no lithium flag set.

@@

We downloaded\footnote{Via \url{datacentral.org.au/services/download},
using the \texttt{sobject\_id} identifiers.} the GALAH DR3 spectra for
all 107 entries.  We then measured the lithium equivalent widths using
the \texttt{specutils} package \citep{specutils_v1pt1}.  We did this
by first focusing in on a 15\AA\ window centered on the 6707.835\AA\ 
lithium doublet.  We then continuum normalized the spectra using a
third-degree Chebyshev series, while excluding any regions that showed
absorption lines ({\it e.g.}, Fe\,\textsc{I} is present at
6703.58\,\AA and 6705.10\,\AA).  We proceeding by fitting a Gaussian
to the continuum normalized spectra, considering only a $\pm 1\,$\AA\ 
window centered on the Li doublet.  The equivalent widths were then
evaluated by numerically integrating the fitted model over the same
window.  Our approach therefore includes the Fe\,6707.44 blend in the
reported Li equivalent widths---which leads to systematic
overestimates of $\approx$10 to 15\,m\AA\ \citep[{\it
e.g.},][]{bouvier_pleiades_lirot_2018}.  Finally, to derive
uncertainties on the EWs, we repeated the procedure twenty times, but
added noise to the spectra, drawn from a normal distribution with a
scale set by the standard deviation of the continuum.  The reported
uncertainties are then drawn from the $84^{\rm th}$ and $16^{\rm th}$
percentiles of the resulting EW distribution.  We verified the overall
scale of our results by comparing the 21 stars that overlapped between
GALAH and Gaia-ESO.  17 of the 21 measurements agreed within
1-$\sigma$.  The remaining four were stars with $T_{\rm
eff}>6200\,{\rm K}$, for which \citet{randich_gaiaeso_2018} reported
Li EW detections ranging from $50$ to $75\,$m\AA, and for which we
found either null results or only slightly significant
($\approx$30m\AA) detections.

@@

One possible explanation is that the observed halo structures are
tidal tails \citep[{\it e.g.},][]{krumholz_star_2019}. The idea is
that as the cluster contracts, stars escape out of the Lagrange points
imposed by the Milky Way's potential, and form leading and lagging
arms due to differential rotation in the Galaxy.  Kinematic evidence
that the core of \cn\ may be undergoing contraction has recently been
presented by \citet{healy_stellar_2020}, which might increase the rate
at which stars escape.  Whether the exact contraction rate, and the
corresponding evaporation rate of the core are sufficient to
explain the formation of the halo has yet to be assessed.  The
presence or lack of epicyclic overdensities in the tails might also
help clarify whether this scenario is happening
\citep{kupper_more_2012}.

A second possibility is that the clusters form in larger and more
dispersed star formation complexes: the stars in the halo need not
have formed in the same ``clump'' as those in the cluster core.  In
other words, the formation environment might be better described as a
giant molecular filament, rather than a giant molecular cloud.  A
sample of such filaments has recently been collected and analyzed by
\citet{zucker_physical_2018}: if the current $\approx$20:1 aspect
ratio of \cn\ were primordial, its structure would match the aspect
ratios of what they term either ``elongated dense core complexes'', or
``bone candidates''.  A more immediate example is the Orion\,A cloud.
\citet{grosschedl_3d_2018} showed using young stellar objects as
tracers that Orion\,A is 90\,pc long, and that it has a dense ``head''
and a lower density ``tail''.

One way to rule between the two possibilities would be to
observationally establish an age-size correlation.  If the
$\approx$100\,Myr halos are produced primarily through evaporation and
expansion, then the younger clusters should be significantly smaller.
A theoretical approach along similar lines would be to perform a
dynamical analysis of whether filaments like Orion\,A, or those in the
\citet{zucker_physical_2018} sample, are at all expected to evolve
into structures resembling \cn.

@@

From their rotation period analysis of \cn,
\citet{fritzewski_rotation_2020} have recently demonstrated that
between 100 and 200\,Myr, young clusters (\cn, Blanco\,1, Psc-Eri,
M\,35, M\,50, and the Pleiades) all show similar rotation period
distributions.  The FGK stars show a ``slow sequence'', a population
of ``rapid rotators'', and in some cases stars in the ``void'' between
the two populations ({\it e.g.,}, Psc-Eri and \cn\ show more stars in
the void than the Pleiades).  The late M dwarfs with detected rotation
periods tend to rotate rapidly ($<3$\,days), though there may also be
late M dwarfs with slow ($>20$\,days) rotation periods.  Although
these late M dwarfs are beyond the magnitude limit of our analysis,
they have been explored in depth for \cn\ by both
\citet{Irwin_NGC2516_2007} and \citet{fritzewski_rotation_2020}.
While a detailed comparison of our results with theirs is beyond the
scope of this work, we emphasize that the main contribution of our
study beyond these previous analyses our sensitivity
across both the core and halo of \cn, and a
corresponding expansion in the number of stars that can be analyzed.
In the following section, we therefore focus on the origins of the
features in the rotation-color diagram, and whether our expanded
rotation sample can provide any new insights.

@@

The other features of the rotation-color diagram are indications that
magnetic braking is not always the most important effect.  The
question of why the rapid rotators exist, for instance, has not yet
been definitively answered.  Models that introduce variable
core-envelope decoupling timescales and pre-MS disk lifetimes can
reproduce each type of population
\citep{Irwin_NGC2516_2007,gallet_improved_2013,gallet_improved_2015},
but the question then shifts to why these decoupling timescales should
differ for stars that are otherwise identical.  Another possible
explanation invokes a spontaneous and random change of magnetic field
topology that causes the dynamo to couple more strongly with the
stellar wind \citep{brown_metastable_2014}.  Other processes could be
external to the star entirely: \citet{qureshi_signature_2018} for
instance have reported that mergers between giant planets and the star
could explain a non-neglible fraction of the rapid rotators.

A separate hypothesis that we can test using the existing data is the
idea that rapid rotation can be explained through binarity.  
Correlations between rapid rotation and binarity in both young
clusters and the field have preivously been noted
\citep{meibom_effect_2007,stauffer_rotation_2016,simonian_rapid_2019,gillen_ngts_2020}
Related effects could include tidal synchronization for the shortest
period binaries, or alternatively pre-MS magnetic locking between the
disk and star \citep[{\it
e.g.},][]{koenigl_disk_1991,long_locking_2005}.

@@

The lithium-rotation correlation (Figure~\ref{fig:lithiumrot}) carries
over beyond just equivalent widths, and to the actual abundance of
lithium in the photosphere, A(Li).  To our knowledge, the first
demonstration of this was by \citet{soderblom_evolution_1993}, who
calculated the relevant LTE curves of growth for the Li\,6708\,\AA\
doublet, and used them to convert equivalent widths to abundances.
Non-LTE corrections affect the abundances by $<0.5\,{\rm dex}$ and do
not change the overall relationships between Li abundance and stellar
temperature or rotation \citep{carlsson_1994,lind_departures_2009}.

More recent studies of lithium and rotation have expanded these
results in the Pleiades, the Psc-Eri stream, and M\,35, among other
clusters
\citep{bouvier_pleiades_lirot_2018,arancibia_2020,jeffries_m35_li_2020,hawkins_2020}.
In \cn, \citet{jeffries_rotation_1998} previously observed a similar
trend that that seen in Figure~\ref{fig:lithiumrot} when analyzing 24
stars in the core of \cn.  Our concatenation of the
\citet{randich_gaiaeso_2018} Gaia-ESO results with the GALAH-DR3
spectra represent a significant expanasion in volume and color range
from this earlier analysis.

>>>>

Figure~\ref{fig:hr} shows the HR diagram in the space of absolute Gaia
$G$ magnitude as a function of observed $\bpmrp$ color.  Magnitudes
are from Gaia EDR3; we performed the Gaia DR2 to EDR3 cross-match
using the pre-computed \texttt{gaiaedr3.dr2\_neighbourhood} table
available at the Gaia archive, and required the closest (proper motion
and epoch-corrected) source to be the single match.  Spectral types
are interpolated from the \citet{pecaut_mamajek_2013} table, after
accounting for reddening as described
below.\footnote{\url{http://www.pas.rochester.edu/~emamajek/EEM_dwarf_UBVIJHK_colors_Teff.txt},
version 2021.03.02.} Stars in the core appear consistent with having a
fixed age and metallicity, and varying mass.  The halo stars show a
similar sequence, though with greater scatter.  One possible
explanation for the additional scatter is that the halo is more
contaminated by field stars.  For instance, $\approx5$ red giants in
the halo must be field interlopers, because their isochronally implied
ages would be inconsistent with that of the cluster.

@@

\begin{figure*}[t]
	\begin{center}
		\leavevmode
		\subfloat{
			\includegraphics[width=0.72\textwidth]{f3a.pdf}
		}

    \vspace{-0.7cm}
		\subfloat{
			\includegraphics[width=0.72\textwidth]{f3b.pdf}
		}
	\end{center}
	\vspace{-0.7cm}
  \caption{ {\bf TESS rotations periods and dereddened Gaia colors
  across the core and halo of \cn.} Sets $\mathcal{A}$ ({\it top})
  and $\mathcal{B}$ ({\it bottom}) undergo successive stages of
  automated cleaning (see Section~\ref{subsubsec:cluster}).  The Pleiades
  \citep[125\,Myr;][]{rebull_rotation_2016a} and Praesepe
  \citep[650\,Myr;][]{douglas_poking_2017} are shown for reference.
  Stars in the core and the halo of NGC\,2516 overlap with the
  Pleiades at $\bpmrpo$$<$1.2.  From $1.2<\bpmrpo<1.7$, some NGC\,2516
  members have longer rotation periods than Pleiades members. Rotation
  periods for comparison field stars are shown in
  Appendix~\ref{app:compstar}.
  \label{fig:rot}
	}
\end{figure*}
\begin{figure*}[t]
	\begin{center}
		\leavevmode
		\subfloat{
			\includegraphics[width=0.72\textwidth]{f3c.pdf}
		}

    \vspace{-0.5cm}
		\subfloat{
			\includegraphics[width=0.72\textwidth]{f3d.pdf}
		}
	\end{center}
	\vspace{-0.7cm}
  \caption{ {\bf Binarity indicators for the rotators in
  Figure~\ref{fig:rot}.} Sets $\mathcal{A}$ ({\it top}) and
  $\mathcal{B}$ ({\it bottom}) undergo the same successive stages of
  cleaning.  Astrometric binaries ($\mathrm{RUWE}>1.2$; 11\% of stars)
  appear in red; photometric binaries ($>$0.3\,mag from an empirical
  isochrone; 20\% of stars) are overplotted in orange.  
  \label{fig:binarity}
	}
\end{figure*}

We began our rotation analysis by considering all \nkinematic\
candidate members of \cn.  For each source, we retrieved all available
CDIPS light curves, on a per-sector basis.  This yielded \ncdips\
stars with at least one sector from a CDIPS light curve.  Each of
these stars is brighter than the CDIPS magnitude limit of
$G_\mathrm{RP}=16$\,mag.  \nkinematicrpltsixteen\ of the stars from the \cn\
source list had $G_\mathrm{RP}<16$.  The difference (65 stars) is due to 35 stars
from \citet{meingast_2021} which were not available at the time of the
CDIPS reductions, and 30 stars falling on the TESS chip gaps.  At the
distance of \cn, the $G_\mathrm{RP}=16$\,mag cutoff imposed during the CDIPS
processing corresponds roughly to $\bpmrpo$ of 2.2, or a spectral type
of $\approx$M2V.

We removed systematic trends common to the light curves on each
CCD in each individual sector, and stitched together the resulting
light curves before searching for rotation-induced periodicity.
Details regarding our detrending approach are presented in
Appendix~\ref{app:detrending}.  After detrending, we proceeded with a
few cleaning steps: we masked 0.7 days at the beginning and end of
each spacecraft orbit, and ran a sliding standard-deviation rejection
window over the light curve, which removed any outlying points within
$\pm3\times$MAD of the median in each window.

We then measured the rotation period from the resulting light curve
using the periodogram implementations in \texttt{astrobase}, namely
the traditional Lomb-Scargle, and the Stellingwerf phase-dispersion
minimization algorithm
\citep{lomb_1976,stellingwerf_period_1978,scargle_studies_1982,stellingwerf_period_2011,bhatti_astrobase_2018}.
We used the CDIPS aperture radius that, based on theoretical
expectations, was expected to give the optimal balance between light
from the target star and the sky background \citep{Sullivan_2015}.  This
typically resulted in a circular aperture radius of either 1 or 1.5 pixels.
We recorded the top five periodogram peaks from each method, and their
corresponding powers.  Finally, as a check on crowding, we recorded
the number of stars within the aperture with greater
brightness than the target star, and with brightness within 1.25 and
2.5 TESS magnitudes of the target star.

@@

\begin{figure*}[t]
	\begin{center}
		\leavevmode
			\includegraphics[width=0.9\textwidth]{f5a.pdf}
	\end{center}
	\vspace{-0.72cm}
  \caption{ {\bf Lithium in the core and halo of NGC\,2516.}
  Equivalent width of the 6708\,\AA\ doublet is shown versus
  dereddened color for all kinematic \cn\ members with Gaia-ESO
  ($N=459$) or GALAH ($N=107$) spectra available.  The GALAH spectra
  comprise slightly over half of the halo stars, due to the
  non-targeted selection function of that survey.  Field stars (gray
  points) are stars with Kepler planets
  \citep{berger_identifying_2018}.  Stars in the Pleiades (125\,Myr)
  with Li detections are from \citet{bouvier_pleiades_lirot_2018};
  Praesepe detections (650\,Myr) are shown in yellow
  \citep{soderblom_praesepe_1993}.  Points with ${\rm
  EW}\approx0\,$m\AA\ are non-detections; all stars shown
  later than $\approx$M0 do not have significant lithium.
  The Li measurements for both core and halo stars in \cn\ are
  consistent with a near-Pleiades age.
  \label{fig:lithiumcorehalo}
  }
\end{figure*}
\begin{figure*}[t]
	\begin{center}
		\leavevmode
		\includegraphics[width=0.9\textwidth]{f5b.pdf}
	\end{center}
	\vspace{-0.7cm}
  \caption{ {\bf Lithium and rotation in NGC\,2516.} Points show all
  kinematic \cn\ members with TESS rotation periods in the Gaia-ESO or
  GALAH surveys, and are colored by equivalent width.  Points with
  reported equivalent widths below 20\,m\AA\ (including
  non-detections) are shown with gray crosses.  The lithium-rotation
  correlation for the K-dwarfs has been observed in other young
  clusters, and is discussed in Section~\ref{subsec:lithium}.
	\label{fig:lithiumrot}
	}
\end{figure*}

@@

The Gaia-ESO collaboration chose candidate \cn\ members to observe
with GIRAFFE and UVES based on previously reported literature members
and publicly available photometry.  Since the existence of the \cn\
halo was not known at the time of target selection, very few halo
stars are in the sample.  For \cn, \citet{randich_gaiaeso_2018}
reported stellar parameters (including lithium equivalent widths and
metallicities) for 796 stars that they considered possible \cn\
members.  Cross-matching these against our kinematic list of
\nkinematic\ candidate members by position and imposing a 0$\farcs$5
maximum separation limit yielded 459 kinematic members with available
spectra, 417 in the core and 42 in the halo of \cn.  The lop-sided
ratio is due to the Gaia-ESO selection function.  An inspection of the
effective temperature {\it vs{.}} lithium equivalent width based on
their Table~2 showed the surprising feature that stars with $T_{\rm
eff}\lesssim 4000\,{\rm K}$ tend to have non-zero equivalent widths,
despite the expectation that their lithium should be fully depleted.
Whether these measurements were significant was not clear, and so we
opted to download the spectra from the ESO archive to re-measure
equivalent widths.

The GALAH DR3 target selection is discussed in depth by
\citet{buder_galah_2020}.  The relevant aspects for our analysis are
that the ``main'' survey targeted $12<V<14$ stars at
$\delta<+10^\circ$, provided the stars were at least ten degrees from
the galactic plane.  The survey is spatially
inhomogeneous.\footnote{See the footprint at
\url{https://www.galah-survey.org/news/announcing-galah-dr3}} Special
targeting for stars in the TESS southern continuous viewing zone, and
for known open cluster members was also performed.  We identified the
candidate \cn\ members for which spectra had been obtained by
searching the \texttt{GALAH\_DR3\_main\_allstar\_v1} catalog, after
excluding stars with the stellar parameter bit flags 1, 2, 3.  This
excludes spectra with unreliable broadening, low S/N, and unreliable
wavelength solutions (see Table~6 of \citealt{buder_galah_2020}).
Since our main focus is measuring equivalent widths for the
Li\,6708\,\AA\ doublet, these were the most relevant bit flags.  Of
our \nkinematic\ candidate \cn\ members, 107 had spectra in GALAH DR3.
51 were in the ``core''; 56 were in the ``halo''.  78 had ``finite''
lithium, i.e., no lithium flag set.
We downloaded\footnote{Via \url{datacentral.org.au/services/download},
using the \texttt{sobject\_id} identifiers.} the GALAH DR3 spectra for
all 107 entries.

@@

We measured the lithium equivalent widths using the \texttt{specutils}
package \citep{specutils_v1pt1}.  We did this by first focusing in on
a 15\AA\ window centered on the 6707.835\AA\ lithium doublet.  We then
continuum normalized the spectra using a third-degree Chebyshev
series, while excluding any regions that showed absorption lines ({\it
e.g.}, Fe\,\textsc{I} is present at 6703.58\,\AA and 6705.10\,\AA).
We proceeding by fitting a Gaussian to the continuum normalized
spectra, considering only a $\pm 1\,$\AA\ window centered on the Li
doublet.  The equivalent widths were then evaluated by numerically
integrating the fitted model over the same window.  Our approach
therefore includes the Fe\,6707.44 blend in the reported Li equivalent
widths---which leads to systematic overestimates of $\approx$10 to
15\,m\AA\ \citep[{\it e.g.},][]{bouvier_pleiades_lirot_2018}.
Finally, to derive uncertainties on the EWs, we repeated the procedure
twenty times, but added noise to the spectra, drawn from a normal
distribution with a scale set by the standard deviation of the
continuum.  The reported uncertainties are then drawn from the
$84^{\rm th}$ and $16^{\rm th}$ percentiles of the resulting EW
distribution.  We verified the overall scale of our results by
comparing our measurements with those of \citet{randich_gaiaeso_2018}.
Most stars with $T_{\rm eff} > 4500\,{\rm K}$ followed a 1-1 relation.
For cooler stars, our EW measurements are systematically lower, as
expected given the depletion timescales for stars at these ages
\citep[{\it e.g.},][]{soderblom_ages_2014}.  We visually verified from
the spectra that, for the most part, we prefer our measurements.

% the 21 stars that overlapped between GALAH and Gaia-ESO.  17
% of the 21 measurements agreed within 1-$\sigma$.  The remaining four
% were stars with $T_{\rm eff}>6200\,{\rm K}$, for which
% \citet{randich_gaiaeso_2018} reported Li EW detections ranging from
% $50$ to $75\,$m\AA, and for which we found either null results or only
% slightly significant ($\approx$30m\AA) detections.

@@

The most likely explanation seems to be that the observed halo
structures are tidal tails \citep[{\it
e.g.},][]{chumak_tails_2006,krumholz_star_2019}. The idea is that
stars near the cluster's tidal radius escape slowly due to the
galactic tide, and subsequently form leading and lagging arms due to
differential rotation in the Galaxy.  The contraction rate of the
cluster can affect this process.  Whether the exact contraction rate
of \cn\ \citep{healy_stellar_2020}, and the corresponding evaporation
rate of the core are sufficient to explain the formation of the halo
has yet to be assessed. 
% The presence or lack of epicyclic
% overdensities in the tails might also help clarify whether this
% scenario is happening \citep{kupper_more_2012}.

A second possibility is that the clusters form in larger and more
dispersed star formation complexes: the stars in the halo need not
have formed in the same ``clump'' as those in the cluster core.  In
other words, the formation environment might be best described as a
giant molecular filament, rather than a giant molecular cloud.  A
sample of such filaments has recently been collected and analyzed by
\citet{zucker_physical_2018}: if the current $\approx$20:1 aspect
ratio of \cn\ were primordial, its structure would match the aspect
ratios of what they term either ``elongated dense core complexes'', or
``bone candidates''.  A more immediate example is the Orion\,A cloud.
\citet{grosschedl_3d_2018} showed using young stellar objects as
tracers that Orion\,A is 90\,pc long, and that it has a dense ``head''
and a lower density ``tail''.

Any differences in the stellar mass function between the core and halo
could be informative, since the tidal tail explanation predicts that
the mean tail star is expected to be less massive than the mean core
star \citep[{\it e.g.},][]{chumak_tails_2006}.
Figure~\ref{fig:massfunction} shows the histogram of $A_G$-corrected
absolute magnitudes, with the corresponding
\citet{pecaut_mamajek_2013} mass estimates.  The stars closest
to the core do
indeed seem to have a greater proportion of the $\gtrsim$0.8$M_\odot$
stars, while the tail stars have a greater proportion of
$\lesssim$0.4$M_\odot$ stars.  The mean star in the inner separation bin has
$({\rm M}_G)_0=7.08$ ($\approx$0.69$M_\odot$), while in the outer bin
the mean star has $({\rm M}_G)_0=7.61$ ($\approx$0.63$M_\odot$).
While the difference does appear significant given the number of
stars involved, the theoretical implications are not entirely clear,
since the ``molecular filament'' explanation might be able to
accommodate mass differences between the core and tail through density
gradients across the initial filament.

One possible approach for distinguishing the two explanations could be
to observationally establish an age-size correlation.  If the
$\approx$100\,Myr halos are produced primarily through evaporation and
expansion, then the younger clusters should be significantly smaller
\citep{chumak_tails_2006,chumak_realOCs_2006}.  A theoretical approach
along complementary lines would be to perform a dynamical analysis of
whether filaments like Orion\,A, or those in the
\citet{zucker_physical_2018} sample are at all expected to evolve into
structures resembling \cn.

\begin{figure*}[t]
	\begin{center}
		\leavevmode
		\includegraphics[width=0.7\textwidth]{f15.pdf}
	\end{center}
	\vspace{-0.7cm}
	\caption{ {\bf Absolute magnitudes (and masses) for the inner and
			outermost stars in \cn.}
		The \nkinematic\ candidate \cn\ members were split into three groups
		according to their distance from the cluster center.
		The inner and outermost groups, each containing 1,099 stars, are shown to
		provide the clearest picture with respect to expectations for mass
		segregation.
		The mean halo star does have lower mass, consistent with expectations
		for tidal tails.
		\label{fig:massfunction}
	}
\end{figure*}

@@

From their rotation period analysis of \cn,
\citet{fritzewski_rotation_2020} have recently demonstrated that
between 100 and 200\,Myr, young clusters (\cn, Blanco\,1, Psc-Eri,
M\,35, M\,50, and the Pleiades) all show similar rotation period
distributions.  The FGK stars show a ``slow sequence'', a population
of ``rapid rotators'', and in some cases stars in the ``void'' between
the two populations ({\it e.g.,}, Psc-Eri and \cn\ show more stars in
the void than the Pleiades).  The late M dwarfs with detected rotation
periods tend to rotate rapidly ($<3$\,days), though there may also be
late M dwarfs with slow ($>20$\,days) rotation periods \citep{stauffer_rotation_2016}.  Although
these late M dwarfs are beyond the magnitude limit of our analysis,
they have been explored in depth for \cn\ by both
\citet{Irwin_NGC2516_2007} and \citet{fritzewski_rotation_2020}.
While a detailed comparison of our results with theirs is beyond the
scope of this work, we emphasize that the main contribution of our
study beyond these previous analyses our sensitivity
across both the core and halo of \cn, and a
corresponding expansion in the number of stars that can be analyzed.
In the following section, we therefore focus on the origins of the
features in the rotation-color diagram, and whether our expanded
rotation sample can provide any new insights.

@@

The other features of the rotation-color diagram are indications that
magnetic braking is not always the most important effect.  The
question of why the rapid rotators exist, for instance, requires
additional physics.  Models that introduce variable
core-envelope decoupling timescales and pre-MS disk lifetimes can
reproduce each type of population
\citep{Irwin_NGC2516_2007,gallet_improved_2013,gallet_improved_2015},
but the question then shifts to why these timescales should
differ for stars that are otherwise identical.  Another possible
explanation invokes a spontaneous and random change of magnetic field
topology that causes the dynamo to couple more strongly with the
stellar wind \citep{brown_metastable_2014}.  Other processes could be
external to the star entirely: \citet{qureshi_signature_2018} for
instance have reported that mergers between giant planets and the star
could explain a non-neglible fraction of the rapid rotators.

A separate hypothesis that we can test using the existing data is the
idea that rapid rotation can be explained through binarity.  
Correlations between rapid rotation and binarity in both young
clusters and the field have preivously been noted
\citep{meibom_effect_2007,stauffer_rotation_2016,simonian_rapid_2019,gillen_ngts_2020}.
Related effects could include tidal synchronization for the shortest
period binaries, or alternatively pre-MS magnetic locking between the
disk and star \citep[{\it
e.g.},][]{koenigl_disk_1991,long_locking_2005}.

@@

The lithium-rotation correlation (Figure~\ref{fig:lithiumrot}) carries
over beyond just equivalent widths, and to the actual abundance of
lithium in the photosphere, A(Li).  To our knowledge, the first
demonstration of this was by \citet{soderblom_evolution_1993}, who
calculated the relevant LTE curves of growth for the Li\,6708\,\AA\
doublet, and used them to convert equivalent widths to abundances.
Non-LTE corrections do
not change the overall relationships between Li abundance and stellar
temperature or rotation \citep{carlsson_1994,lind_departures_2009}.

More recent studies of lithium and rotation have expanded these
results in the Pleiades, the Psc-Eri stream, and M\,35, among other
clusters
\citep{bouvier_pleiades_lirot_2018,arancibia_2020,jeffries_m35_li_2020,hawkins_2020}.
In \cn, \citet{jeffries_rotation_1998} previously observed a similar
trend that that seen in Figure~\ref{fig:lithiumrot} when analyzing 24
stars in the core of \cn.  Our concatenation of the
Gaia-ESO and GALAH-DR3
spectra represents a significant expanasion in volume and color range
from this earlier analysis.

What causes the correlation between lithium abundance and rotation at
fixed stellar mass?  Hints likely lie in shared correlations between lithium
abundance, rotation rate, radius inflation, internal mixing, and
magnetic field strength
\citep{chabrier_evolution_2007,somers_measurement_2017,jeffries_m35_li_2020}.
For instance, one interpretation of the lithium-rotation correlation
based on pure hydrodynamics is that the rapid rotation leads to less
efficient convective penetration (``overshoot''), which then lowers the mixing
efficiency at the convective-radiative boundary
\citep{baraffe_lithium_2017}.  The magnetic field itself may even be
sufficient to inhibit the convection \citep{ventura_Li_B_1998}.
Alternatively, the star's temperature and density profiles could be at fault: rapid rotators
have the strongest magnetic fields and the most spotted surfaces.
These spots lower the photospheric flux, and drive the stellar radius
to expand while lowering the core temperatures, which in turn slows
the rate of lithium-destroying proton capture reactions
\citep{feiden_magnetic_2013,somers_rotation_2015}.
